---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, chunk = TRUE}
data <- read.csv("./data/activity.csv")
head(data)
```

## What is mean total number of steps taken per day?

First, the calculation of steps per day
```{r, chunk = TRUE}
steps_day <- tapply(data$steps, data$date, sum)
```

Making the histogram
```{r, chunk = TRUE}
hist(steps_day, breaks = 10, xlab = "Steps by day", 
     main = "Frequency of steps taken by day", col = "red")
```

```{r, chunk = TRUE}
mean_steps <- mean(steps_day, na.rm = TRUE)
median_steps <- median(steps_day, na.rm = TRUE)
```
The mean number of steps is `r mean_steps` and the median is `r median_steps`


## What is the average daily activity pattern?

Separating by interval
```{r, chunk = TRUE}
avg_interval <- aggregate(steps ~ interval,data = data, mean, na.rm = TRUE)
```

Making the plot
```{r, chunk = TRUE}
library(ggplot2)
g <- ggplot(avg_interval, aes(interval, steps))
g + geom_line(size = 1.5, color = "blue") + 
        labs(x = "Interval", y = "Average steps", 
             title = "Average steps per interval")
```

```{r, chunk = TRUE}
max <- max(avg_interval$steps)
row_max <- which(avg_interval$steps == max)
max_steps <- avg_interval[row_max,1]

```
The interval with the highest average of steps is **`r max_steps`**

## Imputing missing values

```{r, chunk = TRUE}
number_na <- which(is.na(data) == TRUE)
len <- length(number_na)
```
The total numbers of missing values is **`r len`**


Creating the new dataset without the missing values
```{r, chunk = TRUE}
new_data <- data
mean_interval <- tapply(new_data$steps, new_data$interval, mean, na.rm=TRUE)
new_data$steps[number_na] <- mean_interval[as.character(new_data$interval[number_na])]

steps_day2 <- tapply(new_data$steps, new_data$date, sum)
```

Creating the histogram
```{r, chunk = TRUE}
par(mfcol = c(2,1))
hist(steps_day2, breaks = 10, xlab = "Steps by day", col = "green",
          main = "Frequency of steps taken by day (NA removed)")

hist(steps_day, breaks = 10, xlab = "Steps by day", 
     main = "Frequency of steps taken by day (with NA´s)", col = "red")
```

```{r, chunk = TRUE}
mean_steps2 <- mean(steps_day2, na.rm = TRUE)
median_steps2 <- median(steps_day2, na.rm = TRUE)
```
The mean and median number of steps without removing the missing values is `r mean_steps` and `r median_steps`

The mean and median number of steps removing the missing values is `r mean_steps2` and `r median_steps2`

As you can see, the difference is not very significant

## Are there differences in activity patterns between weekdays and weekends?

Creating the new factor variable

```{r, chunk = TRUE}
new_data$date <- as.Date(new_data$date)

library(dplyr)

new_data <- mutate(new_data, 
                   week = ifelse(weekdays(new_data$date) == "sábado" | 
                             weekdays(new_data$date) == "domingo",
                   "weekend", "weekday"))

new_data$week <- as.factor(new_data$week)
```

Making the graphic

```{r, chunk = TRUE}
weekType_avg <- aggregate(steps ~ interval + week, data = new_data, mean)


g2 <- ggplot(weekType_avg, aes(interval, steps, color = week)) + 
        geom_line(size = 1.5) + 
        facet_grid(week~.) + labs(x = "Interval", y = "Average steps",
                                  title = "Mean steps per type of weekday")

g2
```


